---
title: "Rachel Jakielski CS1675 Final Project 01"
output: html_document
---

Collaborators: Bruno Arnabar 

**PART ONE : Exploration**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(ggridges)
```

```{r, base_data}
df<- readr::read_csv("cs_1675_fall2021_finalproject.csv", col_names = TRUE)
df <- df %>% 
      mutate(outcome = ifelse(output < 0.33, 'event', 'non_event'),
      outcome = factor(outcome, levels = c("event", "non_event"))) %>%
      mutate(y = boot::logit(output)) %>% glimpse()
```
```{r, derived_data}
derived_df <- df %>% 
              mutate(x5 = 1 - (x1 + x2 + x3 + x4),
                     w = x2 / (x3 + x4),
                     z = (x1 + x2) / (x5 + x4),
                     t = v1 * v2)
```

## Data to be explored: 
  **Base features:**   
    **> Chemistry variables: x1:x4 **  
    **> Manufacturing process variables: v1:v5, m **  
  ***Derived features: x5, w, z, and t ***  
    **> x5 = 1 - (x1 + x2 + x3 + x4)**  
    **> w = x2 / (x3 + x4)**  
    **> z = (x1 + x2) / (x4 + x5)**  
    **> t=v1*v2**  
  ***Output: output***  
  ***Logit-transformed response: y = boot::logit(output)***  
  
### Check datatypes per feature:  

```{r, base_types}
visdat::vis_dat(df) #checking dataypes per feature
```

### Check the distinct values per base feature:  


```{r, base_distinct}
df %>% subset(select = -output)%>% purrr::map_dbl(n_distinct) #distinct values per feature
```

All base features have roughly the same number of observations, with the exception of m. 
The machine (m) input has relatively few unique values. Visually checking the counts associated with each unique value for m:  

```{r, check_m}
#df %>% count(m) #counts
df%>%
  subset(select = -output)%>% 
  ggplot(mapping=aes(m)) +
  geom_bar() +
  theme_bw()
```


### Visualizing the distributions of chemistry and manufacturing process features: 

```{r, visualize_base}
df %>% 
  select(starts_with('x'), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(starts_with('x'))%>%
  ggplot(mapping = aes(x= value)) +
  geom_histogram(bins = 21) +
  facet_wrap(~name, scales = 'free_x')

df %>% 
  select(starts_with('v'), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(starts_with('v'))%>%
  ggplot(mapping = aes(x= value)) +
  geom_histogram(bins = 21) +
  facet_wrap(~name, scales = 'free_x')
```

### Visualizing the distributions of derived features: 

```{r, visualize_derived}
derived_df %>% 
  select(c(x5,t,w,z)) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(c(x5,t,w,z))%>%
  ggplot(mapping = aes(x= value)) +
  geom_histogram(bins = 21) +
  facet_wrap(~name, scales = 'free_x')
```
### Distribution of the responses: 

```{r, distribution_of_response}
derived_df %>% 
  select(c(y,output)) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(c(y,output))%>%
  ggplot(mapping = aes(x= value)) +
  geom_histogram(bins = 21) +
  facet_wrap(~name, scales = 'free_x')
```

The logit transformed response pushes the output to resemble a normal distribution. 

### Visualizing expanded features per machine: 

```{r, base_by_machine}
derived_df %>% 
  select(c(x5,t,w,z), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(c(x5,t,w,z))%>%
  ggplot(mapping = aes(x= value)) + 
  geom_freqpoly(mapping = aes(color = m, y = stat(density))) + 
  facet_wrap(~name, scales = 'free')
```

### Visualizing output features per machine: 

```{r, output_by_machine}
derived_df %>% 
  select(c(output,y), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(c(output,y))%>%
  ggplot(mapping = aes(x= value)) + 
  geom_freqpoly(mapping = aes(color = m, y = stat(density))) + 
  facet_wrap(~name, scales = 'free')
```

There seems to be some variation on the output extremes (approaching 1.0) per machine. However, the logic transformed response, y, shows no significant variation. 

### Visualizing chemistry & manufacturing process input per by machine: 


```{r, expanded_by_machine}
df %>% 
  select(starts_with('x'), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(starts_with('x'))%>%
  ggplot(mapping = aes(x= value)) + 
  geom_freqpoly(mapping = aes(color = m, y = stat(density))) + 
  facet_wrap(~name, scales = 'free')

df %>% 
  select(starts_with('v'), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(starts_with('v'))%>%
  ggplot(mapping = aes(x= value)) + 
  geom_freqpoly(mapping = aes(color = m, y = stat(density))) + 
  facet_wrap(~name, scales = 'free')

```

No feature varies significantly according to 'm'. 

### Visualizing the summary statistics of base chemistry and manufacturing features: 


```{r, base_distribution}
df %>% 
  select(starts_with('x'), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(starts_with('x'))%>%
  ggplot(mapping = aes(x= m)) + 
  geom_boxplot(mapping = aes( y = value))+
  stat_summary(fun.data = 'mean_se', mapping = aes(y = value), fun.args = list(mult = 2))+
  facet_wrap(~name, scales = 'free')
```

The distributions of each chemistry & manufacturing process feature appear roughly uniform across machines; machine doesn't appear to impact any single variable in particular. 


### Visualizing the summary statistics of the expanded feature set: 

```{r,derived_distribution}
derived_df %>% 
  select(c(x5,t,w,z), m) %>%
  tibble::rowid_to_column() %>%
  pivot_longer(c(x5,t,w,z))%>%
  ggplot(mapping = aes(x= m)) + 
  geom_boxplot(mapping = aes( y = value))+
  stat_summary(fun.data = 'mean_se', mapping = aes(y = value), fun.args = list(mult = 2))+
  facet_wrap(~name, scales = 'free')
```

Features 't' and 'x5' appear to contain high value outliers. However, machine still does not appear to affect any derived feature. 

## Visualizing correlation between inputs 

### Correlation between base & derived features: 

```{r, input_correlation}
derived_df %>%
  select(-m, -output, -outcome) %>%
  cor() %>%
  corrplot::corrplot(method = 'number',type = 'upper')
```

No input has significant correlation (>0.85). However, features x5 and z seem to have a moderately negative correlation with respect to each other. 

# Visualizing the relationship between the logit-transformed response, y, and output with respect to inputs:

### Base Features: 

```{r, x_against_output}
df %>% 
  select(starts_with('x'), -x4, y) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "y")) %>% 
  ggplot(mapping = aes(x = value, y = y)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()

df %>% 
  select(starts_with('x'), -x4, output) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "output")) %>% 
  ggplot(mapping = aes(x = value, y = output)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()
```

```{r, x4_against_output}
df %>% 
  select(x4, y) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "y")) %>% 
  ggplot(mapping = aes(x = value, y = y)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()

df %>% 
  select(x4, output) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "output")) %>% 
  ggplot(mapping = aes(x = value, y = output)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()
```

```{r,v_against_y}
df %>% 
  select(v1,v3,v5, y) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "y")) %>% 
  ggplot(mapping = aes(x = value, y = y)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()


df %>% 
  select(v2,v4, y) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "y")) %>% 
  ggplot(mapping = aes(x = value, y = y)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()
```
```{r,v_against_output}
df %>% 
  select(v1,v3,v5, output) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "output")) %>% 
  ggplot(mapping = aes(x = value, y = output)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()


df %>% 
  select(v2,v4, output) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "output")) %>% 
  ggplot(mapping = aes(x = value, y = output)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()
```

There seems to be little to no trend with respect to 'y' and inputs 'v' and 'x4'. There seem to be *some* relationship between certain values of 'x1', 'x2', and 'x3' that result in a minimized 'y'. 

### Derived Features: 


```{r, t_z_against_y}
derived_df %>% 
  select(t,z, y) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "y")) %>% 
  ggplot(mapping = aes(x = value, y = y)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()

derived_df %>% 
  select(t,z, output) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "output")) %>% 
  ggplot(mapping = aes(x = value, y = output)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()
```
```{r, w_x5__against_y}
derived_df %>% 
  select(w,x5, y) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "y")) %>% 
  ggplot(mapping = aes(x = value, y = y)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()

derived_df %>% 
  select(w,x5, output) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "output")) %>% 
  ggplot(mapping = aes(x = value, y = output)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~name) +
  theme_bw()
```

It does nto appear that there is a relationship between values of 't' that result in a minimized 'y'. There may be a relationship between certain values of 'w', 'x5', and 'z' that result in a minimized 'y'. 


# Visualizing the behavior between the derived binary outcome and the inputs: 

```{r}
derived_df %>% 
  ggplot(mapping = aes(x = outcome)) +
  geom_bar() +
  geom_text(stat = 'count',
            mapping = aes(label = stat(count)),
            color = 'red',
            nudge_y = 7,
            size = 5.5) +
  theme_bw()
```

Although class "event" occurs at a rate of ~50% of class "non event", the imbalance is not substantial; subsampling should not be required. 

```{r, viz_inputs_x}
derived_df %>% 
  select(starts_with('x'),outcome) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "outcome")) %>% 
  mutate(input_id = as.integer( stringr::str_extract(name, "\\d+") )) %>% 
  mutate(input_bin = cut(input_id,
                         breaks = quantile(input_id),
                         include.lowest = TRUE)) %>% 
  filter(input_id < 11) %>% 
  ggplot(mapping = aes(x = value, y = as.factor(input_id))) +
  geom_density_ridges(mapping = aes(fill = outcome),
                      alpha = 0.5) +
  ggthemes::scale_fill_colorblind() +
  theme_bw() +
  theme(strip.text = element_blank())
```

```{r, viz_inputs_v}
derived_df %>% 
  select(starts_with('v'),outcome) %>%
  tibble::rowid_to_column() %>% 
  pivot_longer(!c("rowid", "outcome")) %>% 
  mutate(input_id = as.integer( stringr::str_extract(name, "\\d+") )) %>% 
  mutate(input_bin = cut(input_id,
                         breaks = quantile(input_id),
                         include.lowest = TRUE)) %>% 
  filter(input_id < 11) %>% 
  ggplot(mapping = aes(x = value, y = as.factor(input_id))) +
  geom_density_ridges(mapping = aes(fill = outcome),
                      alpha = 0.5) +
  #facet_wrap(~input_bin, scales = "free_y") +
  ggthemes::scale_fill_colorblind() +
  theme_bw() +
  theme(strip.text = element_blank())
```
