---
title: "Rachel Jakielski CS1675 Final Project 01"
output: html_document
---

***PART TWO: Linear Models***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
library(tidyverse)

library(caret)

library(coefplot)

library(splines)
```


```{r, base_data}
df<- readr::read_csv("cs_1675_fall2021_finalproject.csv", col_names = TRUE)
df <- df %>% 
      mutate(y = boot::logit(output)) 
df <- subset(df, select = -output)

glimpse(df)
```

```{r, derived_data}
derived_df <- df %>% 
              mutate(x5 = 1 - (x1 + x2 + x3 + x4),
                     w = x2 / (x3 + x4),
                     z = (x1 + x2) / (x5 + x4),
                     t = v1 * v2) %>% 
              glimpse()
```

# Base Feature Models: 


```{r, base_linear_additive}
lm_base_additive <- lm(y ~ (.), 
                   data = df)

#coefplot(fit_lm_01)
fit_lm_01_performance <- broom::glance(lm_base_additive)
```

```{r, base_interaction}
lm_base_interaction <- lm(y ~ m*(.), 
                   data = df)
#coefplot(fit_lm_02)
fit_lm_02_performance <- broom::glance(lm_base_interaction)
```

```{r, base_pairwise}
lm_base_pairwise <- lm(y ~ (.)^2, 
                   data = df)
#coefplot(fit_lm_03)
fit_lm_03_performance <- broom::glance(lm_base_pairwise)
```

# Expanded Feature Models: 

```{r, expand_linear_additive}
lm_expanded_additive <- lm(y ~ (.-x4-v1), 
                   data = derived_df)
#coefplot(expanded_lm_01)
expanded_lm_01_performace <- broom::glance(lm_expanded_additive)
```

```{r, expand_interaction}
lm_expanded_interaction <- lm(y ~ m*(.-x4-v1), 
                   data = derived_df)

#coefplot(expanded_lm_02)
expanded_lm_02_performace <- broom::glance(lm_base_interaction)
```

```{r, expanded_pairwise}
lm_expanded_pairwise <- lm(y ~ (.-x4-v1-m)^2, 
                   data = derived_df)

#coefplot(expanded_lm_03)
expanded_lm_03_performace <- broom::glance(lm_expanded_pairwise)
```

# Linear Basis Function Models: 

```{r, lb_01}
lbm_01 <- lm(y ~ x1*(.-x4-v1-w), data = derived_df)
#coefplot(lbm_01)
lbm_01_performace <- broom::glance(lbm_01)
```

```{r, lb_02}
lbm_02 <- lm(y ~ splines::ns(x1,df=4) * (w+z+x2+x3), 
                 data = derived_df)
#coefplot(lbm_02)
lbm_02_performace <- broom::glance(lbm_02)
```

```{r, lb_03}
lbm_03 <- lm(y ~ m*(. -v1 -v2 -v3 -v4 -v5 -x4), 
                 data = derived_df)
#coefplot(lbm_03)
lbm_03_performace <- broom::glance(lbm_03)
```

# Model Analysis: 

```{r, performance_df}
performance_list <- list(fit_lm_01_performance, fit_lm_02_performance, fit_lm_03_performance, expanded_lm_01_performace, expanded_lm_02_performace, expanded_lm_03_performace, lbm_01_performace, lbm_02_performace, lbm_03_performace)

rnames = c("base_1", "base_2", "base_3", "expanded_1", "expanded_2","expanded_3", "basis_1", "basis_2", "basis_3")
      
performance_df <- do.call(rbind.data.frame,performance_list)

performance_df
```


According to AIC and BIC, the best 3 models are: 

1. lbm_02, splines::ns(x1,df=4) * (w+z+x2+x3) (model #8) 

2. lm_expanded_pairwise, y ~ (.-x4-v1-m)^2 (model #6) 

3. lbm_01, y ~ x1*(.-x4-v1-w) (model #7) 
 

```{r}

coefplot(lbm_02)

```

```{r}

coefplot(lm_expanded_pairwise)

```

```{r}

coefplot(lbm_01)

```

# Comparison of coefficient plots for the top 3 models: 

All models indicate x1 and x2 as especially significant. The manufacturing process inputs, v1:v5, behave similarly across models lbm_01 and lm_expanded_pairwise. Interactions with the manufacturing processes have a comparable behavior to these inputs alone. The expanded features, t & z, also show similar behavior across the top 3 models. 

Saving Best Model: 

```{r, save_models}
lbm_02 %>% readr::write_rds("BASIS_2.rds")
```


